Agents and you, a brief summary

### 2015-11-11 methods/agents/summery.tex
added a bunch of extra info about different modal logics

\paragraph{Semantics} Kripke %KRIPKE, *waves arm*.
model/frame $M=\langle S,\pi,R \rangle$

@@

\begin{tabular}{llll}
	% I added the natural collumn, because otherwise I won't remember.
	% Translating something in a different notation is stupid.
	notation & = & definition & natural \\ \toprule
	$R_{\alpha;\beta}$ & = & $R_\alpha \circ R_\beta$ &
	execute from $R_\alpha$ until including $R_\beta$\\
	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
	execute $R_\alpha$ or $R_\beta$\\
	$R_{\alpha*}$ & = & $R_\alpha*$ &
	Execute an arbitrary times \\ \bottomrule
\end{tabular}

@@

	symmetrical &$\forall s,t \in S:(s,t) \in R \Rightarrow (t,s) \in R$ &
	undirected, can go back and forth\\
	euclidiean & $\forall s,t,u \in S:(s,t) \in R \& (s,u) \in R \Rightarrow
	(t,u) \in R$ & \brcell{if s has a relation with t and u, then u and t \\
	have a relation to, you can make triangles.} \\
	serial & $\forall s \in S \exists t \in S (s,t) \in R$&
	\brcell{All states in S are connected to \\ at least one other state in S}\\
	an equivalance relation& R is \emph{reflexive, transitive and 
	symetrical}& \brcell{the states R links to have something in \\ 
	common (not equal?)} \\ \bottomrule
\end{tabular}
\section{APL}
Medhi dastani taught about the agent programming langauge (APL).
\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

>>>>

\paragraph{Semantics} Kripke %KRIPKE, *waves arm*.
model $M=\langle S,\pi,R \rangle$, kripke frame $M=\langle S,\pi,R \rangle$
(truth assignment function is abstracted away)

@@

\begin{tabular}{llll}
	% I added the natural collumn, because otherwise I won't remember.
	% Translating something in a different notation is stupid.
	notation & = & definition & natural \\ \toprule
	$R_{\alpha;\beta}$ & = & $R_\alpha \circ R_\beta$ &
	execute from $R_\alpha$ until including $R_\beta$\\
	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
	execute $R_\alpha$ or $R_\beta$ or both\\
	$R_{\alpha*}$ & = & $R_\alpha*$ &
	Execute an arbitrary times \\ \bottomrule
\end{tabular}

@@

	symmetrical &$\forall s,t \in S:(s,t) \in R \Rightarrow (t,s) \in R$ &
	undirected, can go back and forth\\
	euclidiean & $\forall s,t,u \in S:(s,t) \in R \& (s,u) \in R \Rightarrow
	(t,u) \in R$ & \brcell{if s has a relation with t and u, then u and t \\
	have a relation to, you can make triangles.} \\
	serial & $\forall s \in S \exists t \in S (s,t) \in R$&
	\brcell{All states in S are connected to \\ at least one other state in S}\\
	an equivalance relation& R is \emph{reflexive, transitive and 
	symetrical}& \brcell{the states R links to have something in \\ 
	common (not equal?)} \\ \bottomrule
\end{tabular}
\subsection{Classes of models}
\begin{tabular}{lll}
	Class & properties & axioms \\ \toprule
	$K$	& No constraints & All propositional +
	$\{(\Box\phi \wedge \Box(\phi \to \psi)) \to \Box \phi\}$ \\
	$T$	& reflexive	& K + $\{\Box \phi \to \phi\}$ \\
	$S4$	& reflexive transative	& T + $\{\Box \Box \phi \to \Box \phi\}$ \\
	$S5$	& equivalence & S4 + $\{\neg \Box \phi \to \Box\neg\Box \phi\}$ \\
	$KD$	& serial & K + $\{\neg \Box \bot\}$ \\
	$KD45$	& serial, transitive and euclidean & S5 + $\{\neg \Box \bot\}$ \\
\end{tabular}
Alternative formulation of the 5-axiom:

\begin{prooftree}
	\AxiomC{$\neg \Box \phi \to \Box\neg\Box \phi$}
	\RightLabel{$\Box \phi = \neg\Diamond\neg\phi$}
	\UnaryInfC{$\Diamond \neg \phi \to \Box\Diamond \neg \phi$}
	\RightLabel{Substitute $\neg \phi$ for $\psi$}
	\UnaryInfC{$\Diamond \psi \to \Box\Diamond \psi$}
	\RightLabel{Substitute $\psi$ for $\phi$}
	\UnaryInfC{$\Diamond \phi \to \Box\Diamond \phi$}
\end{prooftree}

\subsection{Deontic logic}
The system $KD$ is also known as $SDL$ (standart deontic logic)
% sdl = standart library as everyone knows.
Deontic logic is the logic of \emph{obligation, prohibition and permission}
or rather: the logic of \emph{ideal vs actual sistuations}

\subsubsection{Rules}
\begin{itemize}
	\item Propositional calculus (including modus ponens)
	\item $(O\phi \wedge O(\phi \to \psi)) \to O \phi$ (K axiom)
	\item $\frac{\phi}{O\phi}$ (necisitation rule)
	\item $\neg O \top $ (D-axiom: obligation is consistent)
	\item $F\phi \leftrightarrow O\neg \phi$ (forbidden is obliged to not)
	\item $P\phi \leftrightarrow \neg F \phi$ (permitted is not forbidden)
\end{itemize}

\subsection{Temporal Logic}
Basic linear time logic (LTL), Time as accisibility relation, its reflexive
and transitive.

Viewed in this way: LTL= S4, $\Box$ stands for `always in the future'.
By convention the present is included in the future.

\subsubsection{Rules}
\begin{itemize}
	\item Propositional calculus (including modus ponens)
	\item $(\Box\phi \wedge \Box(\phi \to \psi)) \to \Box \phi$ (K axiom)
	\item $\frac{\phi}{\Box\phi}$ (necisitation rule)
	\item $\Box \phi \to \phi$ (always implies now)
	\item $\Box \phi \to \Box\Box\phi$ (always implies always)
\end{itemize}

\subsection{Epistemic \& doxastic logic}
For \emph{knowledge} we take the relation $R_i$: reflexive, (transative) and
eculidean, i.e. An equivalance relation.

For \emph{believe} we take the relation $R_i$: serial, transitive and 
euclidean.

\subsubsection{Epistemic rules}
\begin{itemize}
	\item Propositional calculus (including modus ponens)
	\item $(K\phi \wedge K(\phi \to \psi)) \to K \phi$ (K axiom)
	\item $\frac{\phi}{K\phi}$ (necisitation rule)
	\item $K \phi \to \phi$ (knowledge is true)
	\item $K \phi \to KK\phi$ (positive introspection)
	\item $\neg K \phi \to K\neg K\phi$ (negative introspection)
\end{itemize}

\subsubsection{Doxastic rules}
\begin{itemize}
	\item Propositional calculus (including modus ponens)
	\item $(B\phi \wedge B(\phi \to \psi)) \to B \phi$ (B axiom)
	\item $\frac{\phi}{B\phi}$ (necisitation rule)
	\item $\neg B \bot$ (belief is consistent)
	\item $B \phi \to BB\phi$ (positive introspection)
	\item $\neg B \phi \to B\neg B\phi$ (negative introspection)
\end{itemize}
\section{APL}
Medhi dastani taught about the agent programming langauge (APL).
\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\
